Full title of the paper (Capitalized)

### 2023-11-02 MDPI_submition.Rmd
Commit 2023-11-02 06:14

For completeness, we will provide below a brief overview of the clear sky identification algorithm (CSid), along with the site specific thresholds.
To calculate the reference clear sky
$\text{SDR}_\text{CSref}$ we used the $\text{SDR}_\text{Haurwitz}$ derived by
the radiation model of @Haurwitz1945, adjusted for our site with a
factor $a$ (Eq.\ \ref{eq:ahau}), resulted by an iterative optimization process, as described
by @Long2000 and @Reno2016.
The target of
the optimization was the minimization of a function $f(a)$ (Eq.\ \ref{eq:minf}) and
was accomplished with the algorithmic function "optimise", which is an implementation based on the work of @Brent1973, from the library "stats" of the R programming language [@RCT2023]. 
The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 - 2021.
During the optimization, eight simple clear sky radiation models were tested (Daneshyar-Paltridge-Proctor, Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
These models are described in more details by @Reno2012 and evaluated by @Reno2016.
We found, that Haurwitz's model, adjusted with the factor $a = `r signif(alpha,3)`$ yields one of the lowest root mean squared errors (RMSE), 
<!--[....rephrased....]{.comment-start id="3" author="A N" date="2023-07-15T20:32:50Z"}-->
while the procedure, manages to characterize the majority of the data.
<!--[]{.comment-end id="3"}--> 
Thus, our clear sky reference is derived by the Eq.\ \ref{eq:ahau}.

@@

The criteria that were used to identify whether a measurement was taken under clear-sky conditions are presented below.
A data point is flagged as "clear-sky" if all criteria are satisfied, otherwise it is considered to be "cloud-sky".
Each criterion was applied for a running window of $`r nt`$ consecutive one-minute measurements, and the characterization is assigned to the central value of the window.
Each parameter, was calculated both from the observations and the reference clear sky model, for each comparison.
The allowable range of variation is defined by the model-derived value of the parameter multiplied by a factor plus an offset. 
The factors and the offsets were determined empirically, by manual inspecting each filters performance on selected days, and adjusting them accordingly, during an iterative process.

@@

@@

>>>>

For completeness, we will provide below a brief overview of the 
<!--
[....Csid term is referenced above....]{.comment-start id="0" author="A N"
date="2023-07-15T20:00:52Z"} 
-->
clear sky identification algorithm (CSid),
<!--
[]{.comment-end id="0"} 
-->
along with the site specific thresholds.
To calculate the reference clear sky
$\text{SDR}_\text{CSref}$ we used the $\text{SDR}_\text{Haurwitz}$ derived by
the radiation model of @Haurwitz1945, adjusted for our site with a
factor $a$ (Eq.\ \ref{eq:ahau}), resulted by an iterative optimization process, as described
by @Long2000 and @Reno2016.
The target of
the optimization was the minimization of a function $f(a)$ (Eq.\ \ref{eq:minf}) and
was accomplished with the algorithmic function "optimise", which is an implementation based on the work of @Brent1973, from the library "stats" of the R programming language [@RCT2023]. 
The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 - 2021.
During the optimization, eight simple clear sky radiation models were tested (Daneshyar-Paltridge-Proctor, Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
These models are described in more details by @Reno2012 and evaluated by @Reno2016.
We found, that Haurwitz's model, adjusted with the factor $a = `r signif(alpha,3)`$ yields one of the lowest root mean squared errors (RMSE), 
<!--[....rephrased....]{.comment-start id="3" author="A N" date="2023-07-15T20:32:50Z"}-->
while the procedure, manages to characterize the majority of the data.
<!--[]{.comment-end id="3"}--> 
Thus, our clear sky reference is derived by the Eq.\ \ref{eq:ahau}.

@@

The criteria that were used to identify whether a measurement was taken
under clear-sky conditions are presented below.
A data point is flagged
as "clear-sky" if all criteria are satisfied, otherwise it is considered to be "cloud-sky".
Each criterion was applied
for a running window of $`r nt`$ consecutive one-minute measurements, and
the characterization is assigned to the central value of the window.
Each parameter, was calculated both from the observations and the
reference clear sky model, for each comparison.
The allowable range of variation is defined by the
model-derived value of the parameter multiplied by a factor plus an
offset. 
<!--[.... .expanded the description ....]{.comment-start id="4"
author="A N" date="2023-07-15T22:48:13Z"}
-->
The factors and the offsets were
determined empirically, by manual inspecting each filters performance on
selected days, and adjusting them accordingly, during an iterative
process.
<!--[[]{.comment-end id="4"}]{.insertion author="A N"
date="2023-07-15T22:48:13Z"}
-->

@@

```{r echo=F, eval=TRUE}
LT <- data.table(read.csv("./figures/tbl_longterm_trends.dat",
               strip.white = TRUE,
               sep = ";",
               header = TRUE))
LT <- LT[grep("SDR", LT$var), ]
LT$slope.sd           <- NULL
LT$slope.ConfInt_0.95 <- NULL
LT$slope.ConfInt_0.99 <- NULL
LT$var                <- NULL
# LT$N                  <- NULL
LT$Obs                <- NULL
setorder(LT, -slope.stat_sig)
rownames(LT) <- NULL
names(LT)    <- c("Trend [%/year]", 
                  "Trend p-value", 
                  "Sky conditions",
                  "Days",
                  "Statistical signif. [%]")
## no need all are 100%
LTminSSig <- signif(min(LT$`Statistical signif. [%]`), 4)
# LT$`Trend statistical signif. [%]` <- NULL
## numeric precision
LT$`Trend [%/year]` <- signif(LT$`Trend [%/year]`, 3)
LT$`Trend p-value`  <- signif(LT$`Trend p-value`,  3)

## some vars to use later
alltrend   <- signif(LT[`Sky conditions` == "All sky cond.",    `Trend [%/year]`], 2)
cleartrend <- signif(LT[`Sky conditions` == "Clear sky cond.",  `Trend [%/year]`], 2)
cloudtrend <- signif(LT[`Sky conditions` == "Cloudy sky cond.", `Trend [%/year]`], 2)
## rename sky conditions
LT$`Sky conditions` <- sub(" cond.", "", LT$`Sky conditions`)

## order columns
LT1 <- LT[,c(
      "Sky conditions",
      "Trend [%/year]",
      "Trend p-value",
      "Days"
      )]

## order columns
LT2 <- LT[,c(
      "Sky conditions",
      "Trend [%/year]",
      "Statistical signif. [%]",
      "Days"
      )]
```

We calculated the linear SDR trends, from the departures of the mean daily values from the seasonal values, and for the three sky conditions (Table\ \ref{tab:trendtable}).
In Figure\ \ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are very similar and are shown in the Appendix (Figures\ \ref{fig:trendCLEAR} and\  \ref{fig:trendCLOUD}).
We observe a positive trend for all-sky conditions 
($`r alltrend`\,\%/y$), a very close but lower trend for clear-skies ($`r cleartrend`\,\%/y$) and a negative weaker trend for cloud-skies ($`r cloudtrend`\,\%/y$).
In the studied period, there is no significant break or change in the variability of the time series.
Other studies for the European region report a change of the SDR slope, around 1980, a few years before the start of our records [@Wild2021; @Yuan2021; @Ohmura2009]. It is interesting to note, that for the observation period, the trend of the TSI is
$`r format(TSIslopeAll[1,1], scientific = T, digits = 2)`\,\%/y$, and thus we can not attribute any major effect on SDR trend to Solar variability.

<!-- \footnotesize -->
<!-- ```{r trendtable2, echo=F} -->
<!-- panderOptions("table.alignment.default", "center") -->
<!-- pander(LT1, -->
<!--        caption = paste0("\\label{tab:trendtable2}Trends in SDR daily mean for different sky conditions for the period", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ". For all cases the statistical significance is above $", LTminSSig, "\\,\\%$.")) -->
<!-- ``` -->
<!-- \normalsize -->

```{r trendtable, echo=F}
x_latex <- knitr::kable(LT2,
                        format   = "latex",
                        booktabs = T,
                        escape   = T,
                        format.args = list(digits = 4),
                        linesep  = "",
                        align    = "crr",
                        caption  = paste0("\\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), "."))
x_latex %>%
    kable_styling(
        latex_options = c("hold_position")
    )
        # position      =   "float_left"
```

```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap=paste0("Anomaly (\\%) of the daily all-sky SDR, relative to climatological values for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ". The black line shows the long term linear trend.")}
knitr::include_graphics("./images/LongtermTrends-2.png")
```

Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} on Appendix), shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) of the anomalies can reveal different structures within all the three sky conditions [@Regier2019]. 
In the cases of all-sky and clear-sky (Figures \ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three macroscopic periods.
A downward part from the start until about 2005, an relative steady part until about 2016 and finally, a steep upward part until the present.
For cloud-sky (Figure\ \ref{fig:cusummonth-3}), we have a different sequence, it begins with a relative steady part until 1997, followed by a upward part until 2005, a long decline until 2020, with a small positive slope until the present.
For a uniform trend, we would expect the CUSUMs of the anomalies to have a symmetric 'V' shape.
This would indicate that the anomalies are evenly distributed around the climatological mean, and for a positive uniform trend, the first half to be below and the other half above the climatological mean. 
In our case, there is more complex evolution of the phenomenon.
Another distinct feature of the CUSUMs, is the different pattern of the cloudy-sky dataset which peaks around the middle of the period.
Although, it seems to exist a complementary relation to the CUSUMs of the clear- and all-sky cases, we can not assert that clouds are the main drive for this relation, due to the great difference in the number of observational data between the datasets (Table\ \ref{tab:trendtable}).

```{r cusummonth, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
knitr::include_graphics(c(
    "./images/CumulativeMonthlyCuSum-1.png",    # ALL sky
    "./images/CumulativeMonthlyCuSum-5.png",    # CLEAR sky
    "./images/CumulativeMonthlyCuSum-9.png"))   # Cloud sky
```

In order to investigate more, the features within the trends, we created another set of CUSUMs plots by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure\ \ref{fig:cusumnotrendmonthly}).
With this approach periods when the CUSUMs diverge from zero can be interpreted as systematic variation of SDR from the climatological mean. When the CUSUM is increasing, the added values are above the climatological values of SDR trend and vice versa.
Overall, for all- and clear-sky conditions (Figure\ \ref{fig:cusumnotrendmonthly-1} and\ \ref{fig:cusumnotrendmonthly-2}) we observe periods when the anomalies diverge from the climatological value, each lasting for several years. 
The pattern for the all-sky and clear-sky conditions is very similar, suggesting prevalence in clear skies over Thessaloniki.
It is interesting that in the period 1993 - 2016 the anomalies of all-sky and clear-sky, have a high variability around zero. After 2016, the range of the variability is decreased at about one third of the prior period.
For cloud-sky conditions (Figure\ \ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is dominated by positive CUSUMs suggesting a reduced effect of clouds on SDR.
From 1997 to mid-2000s CUSUMs are increasing, likely due to a continuous decrease in the thickness of clouds, followed by a period of rapid increase (within 3 years) in cloud optical thickness lasting up to 2008. 
The following stable period spans for about 15 years up to 2021 when CUSUMs start increasing again.

```{r cusumnotrendmonthly, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR after removing the long term trend from the monthly values.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
knitr::include_graphics(c(
    "./images/CumulativeMonthlyCuSumNOtrend-1.png",    # ALL sky
    "./images/CumulativeMonthlyCuSumNOtrend-5.png",    # CLEAR sky
    "./images/CumulativeMonthlyCuSumNOtrend-9.png"))   # Cloud sky
```

@@

\startlandscape
```{r, mdpinames, echo=FALSE, results='asis'}
mdpi_journals <- "acoustics, actuators, addictions, admsci, adolescents, aerobiology, aerospace, agriculture, agriengineering, agrochemicals, agronomy, ai, air, algorithms,  jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jvd,  wem, wevj, wind, women, world, youth, zoonoticdis"

knitr::kable(rticles::string_to_table(mdpi_journals, 8),
             col.names = NULL,
             format = "latex",
             booktabs = TRUE, 
             caption = "MDPI journal names.",
             longtable = TRUE)

```
\finishlandscape

### 2023-11-02 MDPI_submition.tex
Commit 2023-11-02 06:14

@@

>>>>

We calculated the linear SDR trends, from the departures of the mean
daily values from the seasonal values, and for the three sky conditions
(Table~\ref{tab:trendtable}). In Figure~\ref{fig:trendALL} we present
only the time series under all-sky conditions; the plots for clear-sky
and cloud-sky conditions, are very similar and are shown in the Appendix
(Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}). We observe a
positive trend for all-sky conditions (\(0.38\,\%/y\)), a very close but
lower trend for clear-skies (\(0.35\,\%/y\)) and a negative weaker trend
for cloud-skies (\(-0.28\,\%/y\)). In the studied period, there is no
significant break or change in the variability of the time series. Other
studies for the European region report a change of the SDR slope, around
1980, a few years before the start of our records
\citep{Wild2021, Yuan2021, Ohmura2009}. It is interesting to note, that
for the observation period, the trend of the TSI is \(-2e-04\,\%/y\),
and thus we can not attribute any major effect on SDR trend to Solar
variability.

\begin{table}[!h]

\caption{\label{tab:trendtable}\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
\centering
\begin{tabular}[t]{crrc}
\toprule
Sky conditions & Trend [\%/year] & Statistical signif. [\%] & Days\\
\midrule
All sky & 0.376 & 100.00 & 10256\\
Clear sky & 0.349 & 100.00 & 10256\\
Cloudy sky & -0.276 & 99.97 & 5067\\
\bottomrule
\end{tabular}
\end{table}

{\centering \includegraphics[width=.70\linewidth]{./images/LongtermTrends-2}

\caption{Anomaly (\%) of the daily all-sky SDR, relative to climatological values for 1993 - 2023. The black line shows the long term linear trend.}\label{fig:trendALL}
\end{figure}

Although the year-to-year variability of the anomalies (Figure
\ref{fig:trendALL} and Figures \ref{fig:trendCLEAR},
\ref{fig:trendCLOUD} on Appendix), shows a rather homogeneous behavior,
plots of the cumulative sums (CUSUM) of the anomalies can reveal
different structures within all the three sky conditions
\citep{Regier2019}. In the cases of all-sky and clear-sky (Figures
\ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three
macroscopic periods. A downward part from the start until about 2005, an
relative steady part until about 2016 and finally, a steep upward part
until the present. For cloud-sky (Figure~\ref{fig:cusummonth-3}), we
have a different sequence, it begins with a relative steady part until
1997, followed by a upward part until 2005, a long decline until 2020,
with a small positive slope until the present. For a uniform trend, we
would expect the CUSUMs of the anomalies to have a symmetric `V' shape.
This would indicate that the anomalies are evenly distributed around the
climatological mean, and for a positive uniform trend, the first half to
be below and the other half above the climatological mean. In our case,
there is more complex evolution of the phenomenon. Another distinct
feature of the CUSUMs, is the different pattern of the cloudy-sky
dataset which peaks around the middle of the period. Although, it seems
to exist a complementary relation to the CUSUMs of the clear- and
all-sky cases, we can not assert that clouds are the main drive for this
relation, due to the great difference in the number of observational
data between the datasets (Table~\ref{tab:trendtable}).

{\centering \subfloat[All-skies.\label{fig:cusummonth-1}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-1} }\subfloat[Clear-skies.\label{fig:cusummonth-2}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-5} }\subfloat[Cloud-skies.\label{fig:cusummonth-3}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-9} }

\caption{Running cumulative sum plots of the monthly SDR.}\label{fig:cusummonth}
\end{figure}

In order to investigate more, the features within the trends, we created
another set of CUSUMs plots by subtracting the corresponding long term
trend from the SDR anomaly data, prior to the CUSUM calculation
(Figure~\ref{fig:cusumnotrendmonthly}). With this approach periods when
the CUSUMs diverge from zero can be interpreted as systematic variation
of SDR from the climatological mean. When the CUSUM is increasing, the
added values are above the climatological values of SDR trend and vice
versa. Overall, for all- and clear-sky conditions
(Figure~\ref{fig:cusumnotrendmonthly-1}
and~\ref{fig:cusumnotrendmonthly-2}) we observe periods when the
anomalies diverge from the climatological value, each lasting for
several years. The pattern for the all-sky and clear-sky conditions is
very similar, suggesting prevalence in clear skies over Thessaloniki. It
is interesting that in the period 1993 - 2016 the anomalies of all-sky
and clear-sky, have a high variability around zero. After 2016, the
range of the variability is decreased at about one third of the prior
period. For cloud-sky conditions
(Figure~\ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is
dominated by positive CUSUMs suggesting a reduced effect of clouds on
SDR. From 1997 to mid-2000s CUSUMs are increasing, likely due to a
continuous decrease in the thickness of clouds, followed by a period of
rapid increase (within 3 years) in cloud optical thickness lasting up to
2008. The following stable period spans for about 15 years up to 2021
when CUSUMs start increasing again.

{\centering \subfloat[All-skies.\label{fig:cusumnotrendmonthly-1}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-1} }\subfloat[Clear-skies.\label{fig:cusumnotrendmonthly-2}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-5} }\subfloat[Cloud-skies.\label{fig:cusumnotrendmonthly-3}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-9} }

\caption{Running cumulative sum plots of the monthly SDR after removing the long term trend from the monthly values.}\label{fig:cusumnotrendmonthly}
\end{figure}

@@

\begin{longtable}[t]{llllllll}
\caption{\label{tab:mdpinames}MDPI journal names.}\\
\toprule
acoustics & aerobiology & agronomy & jmse & jor & jsan & wind & \\
actuators & aerospace & ai & jne & journalmedia & jtaer & women & \\
addictions & agriculture & air & jnt & jox & jvd & world & \\
admsci & agriengineering & algorithms & jof & jpm & wem & youth & \\
adolescents & agrochemicals & jmp & joitmc & jrfm & wevj & zoonoticdis & \\
\bottomrule
\end{longtable}
\finishlandscape
